High-Dimensional Regularized Discriminant Analysis

### 2013-12-30 reports/paper-hdrda.Rnw
Add some polish to model-selection discussion. Still need a little more. #11

Thus far, we have presented the \emph{HDRDA} classifier and its properties that
facilitate an efficient calculation of the decision rule. Here, we describe in
an efficient model-selection procedure to select the optimal estimates for the
tuning parameters $\lambda$ and $\gamma$. We provide pseudocode for the model
selection in Algorithm \ref{alg:hdrda_model_selection}. Our goal is to select
the optimal tuning parameter estimates from the Cartesian product of candidate
values, $\{\lambda_g\}_{g=1}^G \times \{\gamma_h\}_{h=1}^H$. We utilize $V$-fold
cross-validation to estimate the conditional error rate for each candidate
tuning-parameter pair and select the pair that attains the minimum error
rate. To calculate the $V$-fold cross-validation, we partition the original
training data into $V$ mutually exclusive and exhaustive folds that have
approximately the same number of observations. Then, for $v = 1, \ldots, V$, we
classify the observations in the $v$th fold by training a classifier on the
remaining $V - 1$ folds for each candidate pair of tuning parameters. We then
calculate the proportion of misclassified observations across the $V$ folds and
select the pair $(\widehat{\lambda}, \widehat{\gamma})$ that minimizes the
considered training error-rate estimator.

@@

    Compute $\bm X_c = \bm M_q \bm D_q \bm U_1^T$

@@

% TODO: Now that the algorithm has been introduced and described, we need to
% discuss the details. Let's remove some of the bloat where we talk about the
% kernel trick, etc. Simply discuss how this technique is fast and cite Hastie's
% book.

A primary contributing factor to the efficiency of Algorithm
\ref{alg:hdrda_model_selection} is our usage of the compact singular value
decomposition (SVD) (sometimes called the reduced or fast SVD), which we
describe here. Rather than computing the eigenvalue decomposition of
$\widehat{\bm \Sigma}$ to obtain $\bm U_1$, we instead obtain $\bm U_1$ by
computing the eigendecomposition of a much smaller $N \times N$ matrix when $p
\gg N$ \cite[Chapter~18.3.5]{Hastie:2008dt}. Recall that the pooled sample
covariance matrix is $\widehat{\bm \Sigma} = \bm X^{T} \bm X$\footnote{TODO:
  $1/N$ is missing. Add this. It will be absorbed into the eigenvalues.}  Then,
because $\bm X = \bm M \bm D^{1/2} \bm U^{T}$ by the SVD, we have the
eigendecomposition $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$, where $\bm M
\in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N
  \times N}^{\ge}$ is a diagonal matrix consisting of the singular values of
$\bm X$, and $\bm U \in \mathbb{R}_{p \times N}$ is orthogonal. Now, $\bm X \bm
X^{T} = \bm M \bm D \bm M^{T}$ is the eigenvalue decomposition of $\bm X \bm
X^{T}$, where $\bm D$ is the matrix of eigenvalues of $\bm X \bm X^{T}$ and the
columns of $\bm M$ are the corresponding eigenvectors. Hence, we obtain $\bm M$
and $\bm D$ from an eigendecomposition of $\bm X \bm X^{T} \in \mathbb{R}_{N
  \times N}$. Next, we compute $\bm U = \bm X^{T} \bm M \bm D^{+/2}$, where $\bm
D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the number
of numerically nonzero eigenvalues present in $\bm D$, by calculating the number
of eigenvalues that exceeds some tolerance value, say, $\epsilon = 1 \times
10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining only the
first $q$ columns of $\bm U$.

As a result of the compact SVD, we need calculate $\bm X_c \bm U_1$ only once
per cross-validation fold, requiring $O(p q N) \approx (p N^2)$ calculations.
Hence, the computational costs of expensive calculations, such as matrix
inverses and determinants, are greatly reduced because they are performed in the
$q$-dimensional subspace. Similarly, we reduce the dimension of the test data
set by calculating $\bm X_c \bm U_1$ once per fold. Conveniently, we see that
the most costly computation involved in $\bm Q_k$ and $\bm W_k^{-1}$ is $\bm X_k
\bm U_1$, which can be extracted from $\bm X_c \bm U_1$. Thus, after the initial
calculation of $\bm X_c \bm U_1$ per cross-validation fold, $\bm Q_k$ requires
$O(TODO)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k \times n_k}$, both
its determinant and inverse requires $O(n_k^3)$ operations. This is especially
advantageous when $K$ is large, such that $n_k$ is much smaller than $N$. Thus,
$\bm W_k^{-1}$ requires $O(TODO)$ operations Also, the inverse and determinant
of the diagonal matrix $\bm \Gamma_k \in \mathbb{R}_{q \times q}$ requires
$O(q)$ operations. Hence, the $| \bm W_k |$ requires $O(q n_k^3)$ operations,
which is efficient when $p >> N$.

The expressions given in Proposition \ref{proposition:hdrda-W_k} also expedite
the selection of $\lambda$ and $\gamma$ via cross-validation because the most
time-consuming matrix operation involved in computing $\bm W_k^{-1}$ and $|\bm
W_k|$ is $\bm X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent
of $\lambda$ and $\gamma$. The subsequent operations in calculating $\bm
W_k^{-1}$ and $|\bm W_K|$ can be simply updated for different pairs of $\lambda$
and $\gamma$.\footnote{TODO: Need to clean up this paragraph so that it flows
  from above. Also, elaborate.}

>>>>

Thus far, we have presented the \emph{HDRDA} classifier and its properties that
facilitate an efficient calculation of the decision rule. Here, we describe an
efficient model-selection procedure along with pseudocode in Algorithm
\ref{alg:hdrda_model_selection} to select the optimal tuning-parameter estimates
from the Cartesian product of candidate values $\{\lambda_g\}_{g=1}^G \times
\{\gamma_h\}_{h=1}^H$. We estimate the $V$-fold cross-validation error rate for
each candidate pair and select $(\widehat{\lambda}, \widehat{\gamma})$ that
attains the minimum error rate. To calculate the $V$-fold cross-validation, we
partition the original training data into $V$ mutually exclusive and exhaustive
folds that have approximately the same number of observations. Then, for $v = 1,
\ldots, V$, we classify the observations in the $v$th fold by training a
classifier on the remaining $V - 1$ folds. We then calculate the
cross-validation error as the proportion of misclassified observations across
the $V$ folds.

@@

    Compute the compact SVD $\bm X_c = \bm M_q \bm D_q \bm U_1^T$

@@

A primary contributing factor to the efficiency of Algorithm
\ref{alg:hdrda_model_selection} is our usage of the compact singular value
decomposition (SVD) (sometimes called the reduced or fast SVD). Rather than
computing the eigenvalue decomposition of $\widehat{\bm \Sigma}$ to obtain $\bm
U_1$, we instead obtain $\bm U_1$ by computing the eigendecomposition of a much
smaller $N \times N$ matrix when $p \gg N$
\cite[Chapter~18.3.5]{Hastie:2008dt}. Recall that the pooled sample covariance
matrix is $\widehat{\bm \Sigma} = \bm X^{T} \bm X$\footnote{TODO: $1/N$ is
  missing. Add this. It will be absorbed into the eigenvalues.}  Then, because
$\bm X = \bm M \bm D^{1/2} \bm U^{T}$ by the SVD, we have the eigendecomposition
$\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$, where $\bm M \in \mathbb{R}_{N
  \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a
diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in
\mathbb{R}_{p \times N}$ is orthogonal. Now, $\bm X \bm X^{T} = \bm M \bm D \bm
M^{T}$ is the eigenvalue decomposition of $\bm X \bm X^{T}$, where $\bm D$ is
the matrix of eigenvalues of $\bm X \bm X^{T}$ and the columns of $\bm M$ are
the corresponding eigenvectors. Hence, we obtain $\bm M$ and $\bm D$ from an
eigendecomposition of $\bm X \bm X^{T} \in \mathbb{R}_{N \times N}$. Next, we
compute $\bm U = \bm X^{T} \bm M \bm D^{+/2}$, where $\bm D^{+/2} = \bm
D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the number of numerically
nonzero eigenvalues present in $\bm D$, by calculating the number of eigenvalues
that exceeds some tolerance value, say, $\epsilon = 1 \times 10^{-6}$. After
calculating $q$, we extract $\bm U_1$ by retaining only the first $q$ columns of
$\bm U$.

As a result of the compact SVD, we need calculate $\bm X_c \bm U_1$ only once
per cross-validation fold, requiring $O(p q N) \approx (p N^2)$ calculations.
Hence, the computational costs of expensive calculations, such as matrix
inverses and determinants, are greatly reduced because they are performed in the
$q$-dimensional subspace. Similarly, we reduce the dimension of the test data
set by calculating $\bm X_{test} \bm U_1$ once per fold. Conveniently, we see
that the most costly computation involved in $\bm Q_k$ and $\bm W_k^{-1}$ is
$\bm X_k \bm U_1$, which can be extracted from $\bm X_c \bm U_1$. Thus, after
the initial calculation of $\bm X_c \bm U_1$ per cross-validation fold, $\bm
Q_k$ requires $O(TODO)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k \times
  n_k}$, both its determinant and inverse require $O(n_k^3)$
operations. Consequently, $\bm W_k^{-1}$ requires $O(TODO)$ operations. Also,
the inverse and determinant of the diagonal matrix $\bm \Gamma_k \in
\mathbb{R}_{q \times q}$ requires $O(q)$ operations. Therefore, the $| \bm W_k
|$ requires $O(q n_k^3)$ operations.

The expressions given in Proposition \ref{proposition:hdrda-W_k} also expedite
the selection of $\lambda$ and $\gamma$ via cross-validation because the most
time-consuming matrix operation involved in computing $\bm W_k^{-1}$ and $|\bm
W_k|$ is $\bm X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent
of $\lambda$ and $\gamma$. The subsequent operations in calculating $\bm
W_k^{-1}$ and $|\bm W_k|$ can be simply updated for different pairs of $\lambda$
and $\gamma$.
